---
title: "BANA 7042_ProjectPart2"
output:
  pdf_document: default
  html_notebook: default
---
#FLIGHT LANDING DISTANCE ANALYSIS - Logistic Regression
This is in continuation of our part1 study but here we will discard the distance variable and work on 2 new binary variables. The dataset is the cleansed dataset we generated during part1 after deleting abnormalities.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(psych)
library(dplyr)
library(funModeling)
library(ggplot2)
```
```{r, warning=FALSE, message=FALSE}
Data <- read.csv("FAA_log.csv", header=TRUE)
str(Data)
```
###Step 1
```{r}
###Adding 2 new binary variables
Data$long.landing <- ifelse(Data$distance>2500,1,0)
Data$risky.landing <- ifelse(Data$distance>3000,1,0)
#Dropping the distance column
Data <- select(Data, -c(distance))
str(Data)
```
The two new variables have been added and distance variable has been dropped successfully.
###Step 2
```{r, warning=FALSE, message=FALSE}
attach(Data)
pct <- round(table(long.landing)/length(long.landing)*100,1)
labs<-c("No","Yes")
labs<-paste(labs,pct)
labs<-paste(labs,"%",sep="")
pie(table(long.landing),labels=labs,col=rainbow(length(labs)),main="Pie chart of long landing variable")
```
#Step-3 Performing single-factor regression analysis
```{r, message=FALSE, warning=FALSE}
lgmod<-glm(long.landing~., family=binomial, Data)
beta.lmod <- coef(lgmod)
round(beta.lmod,3)
```
###Table to rank the factors on the basis on regression summary statistics
```{r, message=FALSE, warning=FALSE}
data.table::data.table(
                     check.names = FALSE,
                        Variable = c("Aircraft","speed_air","height","pitch",
                                     "no_pasg","speed_ground","duration"),
                                       SizeCoef= c(8.766,1.98,0.422,1.463,-0.073,-0.225,0),
                      OddsR = c(6412.471211,7.242742985,1.525008525,
                                     4.318896803,0.92960083,0.798516219,1),
        Direction = c("positive","positive","positive","positive",
                                     "positive","negative","positive"),
                                                            `p-value` = c(0.009,0.005,0.003,0.167,0.296,0.558,0.98                                     )
)
```
###Step4 - Visualize the association
```{r, message=FALSE, warning=FALSE}
ggplot(Data, aes(x=speed_ground, fill=as.factor(long.landing)))+geom_histogram(position="dodge")+theme(legend.position = "top")
```
The data for speed_ground in respect to long landing = 0 is rightly skewed. 
```{r, message=FALSE, warning=FALSE}
ggplot(Data, aes(x=pitch, fill=as.factor(long.landing)))+geom_histogram(position="dodge")+theme(legend.position = "top")
```
The data for pitch in respect to long landing = 0 is normally distributed.
```{r, message=FALSE, warning=FALSE}
ggplot(Data, aes(x=height, fill=as.factor(long.landing)))+geom_histogram(position="dodge")+theme(legend.position = "top")
```
The data for height in respect to long landing = 0 is normally distributed.
```{r, message=FALSE, warning=FALSE}
ggplot(Data, aes(x=speed_air, fill=as.factor(long.landing)))+geom_histogram(position="dodge")+theme(legend.position = "top")
```
The data for speed_air in respect to long landing = 0 is rightly-skewed like speed_ground.

#Checking the collinearity between speed_ground and speed_air
```{r, message=FALSE, warning=FALSE}
plot(Data$speed_air ~ Data$speed_ground, col = rep(c("red", "blue"), each = 50), pch = 16, xlab = "Ground Speed", 
    ylab = "Air speed")
```
As we know, speed_air variable has around 75% of the missing values and there is a high collinearity between speed_ground and speed_variable. To build a proper model it would be wise to consider only one variable, speed_ground going forward.
#Important factors to consider
1. speed_ground
2. height
3. aircraft
4. pitch

###Step5 Fitting the model
```{r, message=FALSE, warning=FALSE}
#Encoding the aircraft variable
Data<-Data%>%
  mutate(aircraft=ifelse(aircraft=="airbus",0,1))
str(Data)
```

```{r, message=FALSE, warning=FALSE}
#Building the full model
full<-glm(long.landing~., Data, family=binomial)
summary(full)
```
#Fitting the model with the important factors
```{r, message=FALSE, warning=FALSE}
#Building the full model
final<-glm(long.landing~speed_ground+height+pitch+aircraft, Data, family=binomial)
summary(final)
```
We can see that all the selected predictors are significant in terms of p-values but pitch. We will use step function to perform alternate analysis and to determine the good fit model.

#Step6 Forward Variable selection using AIC
```{r, message=FALSE, warning=FALSE}
#Removing speed_air column to remove missing values info
FAA.final<-Data%>%
  select(-speed_air)

null_model<- glm(long.landing ~ 1,data=FAA.final,family=binomial)
full_model <- glm(long.landing ~ .,data=FAA.final,family=binomial)
mAIC<-step(null_model,scope=list(lower=null_model, upper=full_model),trace=0,direction = "forward")
summary(mAIC)
```
The AIC method has also chosen the model with our selected factors, selected in Step 3 after our analysis and comparison. We have fitted the model from only one to all the variables except speed_air and AIC returned us the best model with the smallest AIC value.
Thus, the results are consistent with the final decision of important factors we analysed in step 3 to build the model.

#Step7 Forward Variable Selection using BIC
```{r, message=FALSE, warning=FALSE}
mBIC<-step(null_model,scope=list(lower=null_model, upper=full_model),k=log(nrow(FAA.final)))
summary(mBIC)
```
BIC penalizes free parameters more strongly, unlike the AIC. BIC suggested the best(more closer to true model) with 3 factors(speed_ground, aircraft, height) removing the additional and less significant factor(pitch) while AIC provided the good model with all the 4 factors.

#Step8 Important Inferences to present
1. Best model:
#long landing=-102.95 + 0.93*speed_ground + 5.05*aircraft + 0.23*height
2. speed_air contributing to a lot of missing values, so we based our final analysis using speed_ground as the most pertinent factor.
3. Ignoring speed_air wouldn't affect much on our model quality as a strong collinearity was observed between speed_ground and speed_air.
4. The analysis after performing forward selection using AIC and BIC has provided us with the best model having less predictors,3 and lowest AIC, 65.047.
5. p-values are very less than 0.05; and regression coefficients are positive depicting to be in same direction of the response variable.

##Identifying important factors using the binary data of “risky.landing”
Step9 - Performing the same steps for risky.landing variable
```{r, message=FALSE, warning=FALSE}
attach(Data)
pct <- round(table(risky.landing)/length(risky.landing)*100,1)
labs<-c("No","Yes")
labs<-paste(labs,pct)
labs<-paste(labs,"%",sep="")
pie(table(risky.landing),labels=labs,col=rainbow(length(labs)),main="Pie chart of risky landing variable")
```

```{r, message=FALSE, warning=FALSE}
#Fitting the logistic model - single factor analysis
lgmod2<-glm(risky.landing~., family=binomial, Data)
beta.lmod2 <- coef(lgmod2)
round(beta.lmod2,3)
```

###Table to rank the factors on the basis on regression summary statistics
```{r, message=FALSE, warning=FALSE}
data.table::data.table(
                     check.names = FALSE,
                        Variable = c("speed_air","aircraft","height","pitch",
                                     "no_pasg","speed_ground","duration"),
  SizeCoef = c(1.615,7.217,0.044,-1.328,-0.117,-0.177,
                                     0.002),
                      OddsR = c(5.027887923,1362.395737,1.044982355,
                                     0.265006745,0.889585193,0.837779785,1.002002001),
        Direction = c("positive","positive","positive","negative",
                                     "negative","negative","positive"),
                                                            `p-value` = c(0.0142,0.0174,0.454,0.355,0.229,0.73,0.9,
                                     NA)
)
```

Visualize the association
```{r, message=FALSE, warning=FALSE}
ggplot(Data, aes(x=speed_air, fill=as.factor(risky.landing)))+geom_histogram(position="dodge")+theme(legend.position = "top")
```
```{r, message=FALSE, warning=FALSE}
ggplot(Data, aes(x=speed_ground, fill=as.factor(risky.landing)))+geom_histogram(position="dodge")+theme(legend.position = "top")
```
The speed_ground variable is normally distributed for risky.landing = 0
```{r, message=FALSE, warning=FALSE}
ggplot(Data, aes(x=height, fill=as.factor(risky.landing)))+geom_histogram(position="dodge")+theme(legend.position = "top")
```
#The significant factors till now came to be aircraft, speed_ground and height.

```{r, message=FALSE, warning=FALSE}
#fitting the logistic regression model
full2 <- glm(risky.landing~speed_ground+height+aircraft,family=binomial, Data)
summary(full2)
```
#Performing the forward selection using AIC and BIC criteria
```{r, message=FALSE, warning=FALSE}
null_modelR<- glm(risky.landing ~ 1,data=FAA.final,family=binomial)
full_modelR <- glm(risky.landing ~ .,data=FAA.final,family=binomial)
#AIC
AIC1<-step(null_modelR,scope=list(lower=null_modelR, upper=full_modelR),trace=0,direction = "forward")
summary(AIC1)

```
Using BIC
```{r, message=FALSE, warning=FALSE}
BIC1 <- step(null_modelR,scope=list(lower=null_modelR, upper=full_modelR),k=log(nrow(FAA.final)))
summary(BIC1)
```
#Both the methods AIC and BIC have resulted in a model with 2 most significant factors for the resposne variable, risky landing which are speed_ground and aircraft type.

Step10 - Important inferences to present for Risky landings variable
1. Best model: 
#risky landing=-102.08 + 0.93*speed_ground + 4.02*aircraft
2. speed_air contributing to a lot of missing values, so we based our final analysis using speed_ground as the most pertinent factor.
3. The analysis after performing forward selection using AIC and BIC has provided us with the same result of the best model having less predictors,2 and lowest AIC value, 46.097.
4. p-values are very less than 0.05; and regression coefficients are positive depicting to be in same direction of the response variable.


###Compare the two models built for “long.landing” and “risky.landing”
#Step 11
1. Model to predict long landing response variable uses less predictors than risky landing model
2. AIC value for risky landing model is less than long landing.
3. Below is the table for the comparison
```{r, message=FALSE, warning=FALSE}
#Step11
data.table::data.table(
    check.names = FALSE,
       `Model:` = c("No. of variables", "AIC", "Common parameters"),
   long.landing = c(3, 65.047, 2),
  risky.landing = c(2, 46.097, 2)
)
```
#Step12 ROC Curve
```{r, message=FALSE, warning=FALSE}
#long landing model
mlong <-glm(long.landing~aircraft+speed_ground+height, family=binomial,FAA.final)

#risky landing model
mrisky <-glm(risky.landing~aircraft+speed_ground, family=binomial,FAA.final)
```


```{r, message=FALSE, warning=FALSE}
#Model evaluation based on predictive power
library(ROCR)
pred1 <- prediction(predict(mlong), FAA.final$long.landing)
perf1 <- performance(pred1,"tpr","fpr")

pred2 <- prediction(predict(mrisky),FAA.final$risky.landing)
perf2 <- performance(pred2,"tpr","fpr")

#AUC values
unlist(slot(performance(pred1 , "auc"), "y.values"))
unlist(slot(performance(pred2 , "auc"), "y.values"))
```
```{r, message=FALSE, warning=FALSE}
plot(perf1,col="blue", main = 'ROC Curve', xlab = "1-Specificity", ylab = "Sensitvity", sub = "Threshold = 0.5")

plot(perf2,col="red",add=TRUE)

legend(0.6,0.45, c('long landing','risky landing'),lty=c(1,1),
lwd=c(2,2),col=c('blue','red'))

```
ROC curve for long landing and risky landing model depicts that both have almost same AUC close to 1.

#Step13 Predicting the probabilities
```{r, message=FALSE, warning=FALSE}
#Boeing, duration=200, no_pasg=80, speed_ground=115, speed_air=120, height=40, pitch=4
#Aircraft is encoded 0/1, Boeing =1
values<-data.frame(aircraft=1, duration=200, no_pasg=80, speed_ground=115, speed_air=120, height=40, pitch=4)

#Long.landing(L_P) probability and 95% Confidence Interval
P1 <- predict(mlong, newdata=values, type="response", se=T)
CI_L <- c((P1$fit-
                (1.96*P1$se.fit)),(P1$fit+(1.96*P1$se.fit)))
#Risky.landing(R_P) probability and 95% Confidence Interval
P2 <- predict(mrisky, newdata=values, type="response",se=T)
CI_R <- c((P2$fit-
                (1.96*P2$se.fit)),(P2$fit+(1.96*P2$se.fit)))
```

Below is the summary:
```{r, message=FALSE, warning=FALSE}
data.table::data.table(
                         check.names = FALSE,
                         `Model:` = c("Prob", "SE", "95% CI"),
  L_P = c(1, 5.87e-08,"(0.999,1.000)"),
  R_P = c(0.99978, 0.000441, "(0.998,1.006)")
)
```

#Compare models with different link functions
```{r, message=FALSE, warning=FALSE}
#Step14
mr.probit <- glm(risky.landing~aircraft+speed_ground, family=binomial(link = probit),FAA.final)
mr.cloglog <- glm(risky.landing~aircraft+speed_ground, family=binomial(link=cloglog),FAA.final)
mr.logit <- glm(risky.landing~aircraft+speed_ground, family=binomial,FAA.final)

#Table comparing AIC and Standard error for the 3 models
data.table::data.table(
  Parameters = c("AIC", "SE_ground_speed"),
      Probit = c(45.436, 0.12),
     CLogLog = c(47.443, 0.13),
       Logit = c(46.097, 0.22)
)
```
It can be seen that AIC and standard error are the lowest for Probit link function model.

#Step15 Plotting ROC curve of different models for riskylanding variable
```{r, message=FALSE, warning=FALSE}
#Probit model
pred_p <- prediction(predict(mr.probit), FAA.final$long.landing)
perfp1 <- performance(pred_p,"tpr","fpr")

#Hazard model
pred_c <- prediction(predict(mr.cloglog),FAA.final$risky.landing)
perfp2 <- performance(pred_c,"tpr","fpr")

#Logit model
pred_l <- prediction(predict(mr.logit),FAA.final$risky.landing)
perfp3 <- performance(pred_l,"tpr","fpr")


#AUC values
unlist(slot(performance(pred_p , "auc"), "y.values"))
unlist(slot(performance(pred_c , "auc"), "y.values"))
unlist(slot(performance(pred_l , "auc"), "y.values"))
```
The Area under curve is exactly the same for all the different links' models.
#Plotting the ROC Curve
```{r, message=FALSE, warning=FALSE}
plot(perfp1,col="blue", main = 'ROC Curve', xlab = "1-Specificity", ylab = "Sensitvity", sub = "Threshold = 0.5")

plot(perfp2,col="red",add=TRUE)
plot(perfp3,col="green",add=TRUE)

legend(0.6,0.6, c('probit','hazard','logit'),lty=c(1,1),
lwd=c(2,2),col=c('blue','red','green'))
```

#Step16 Identification of top 5 risky landings
```{r, message=FALSE, warning=FALSE}
logit_p<- predict(mr.logit, type = "response")
probit_p<- predict(mr.probit, type = "response")
cloglog_p <- predict(mr.cloglog, type = "response") #Hazard

#Top 5 for logit link
head(sort(logit_p, decreasing = TRUE),5)
```

```{r, message=FALSE, warning=FALSE}
#Top 5 for probit link
head(sort(probit_p, decreasing = TRUE),5)
```

```{r, message=FALSE, warning=FALSE}
#Top 5 for Hazard link
head(sort(cloglog_p, decreasing = TRUE),5)
```
All the 3 links model have retured different top 5 flights. Row 64 is common in all 3 while row 56 is common among probit and hazard link functions model.

#Step17 Using probit and hazard models to make prediction
```{r, message=FALSE, warning=FALSE}
###Probit model

#Predicted Probability
probit_prob <- predict(mr.probit, newdata=values, type="response",se=T)
#95% Confidence interval
CI_probit <- c((probit_prob$fit-             (1.96*probit_prob$se.fit)),(probit_prob$fit+(1.96*probit_prob$se.fit)))

```
```{r, message=FALSE, warning=FALSE}
###Hazard model
#Predicted Probability
hazard_prob <- predict(mr.cloglog, newdata=values, type="response",se=T)
#95% Confidence interval
CI_hazard <- c((hazard_prob$fit-             (1.96*hazard_prob$se.fit)),(hazard_prob$fit+(1.96*hazard_prob$se.fit)))
```
```{r, message=FALSE, warning=FALSE}
#Comparison of 3 different link function models
data.table::data.table(
            check.names = FALSE,
                  Model = c("Logit", "Probit", "Hazard"),
  Pred_Prob = c(0.999789, 0.999, 1),
                     SE = c(5.87e-08, 3.15e-06, 2.61e-16),
               `95%.CI` = c("(0.998,1.006)", "(0.9999933,1.0000056)", "(1,1)")
)
```

